---
title: "Warehouse Data Analysis"
output: html_document
date: "2026-08-12"
---
This notebook is designed to analyze environmental data collected from a warehouse. The analysis will include data cleaning, visualization, and statistical modeling to understand the environmental conditions within the warehouse.

# Setup 

## Load Packages 
```{r}
# Load packages
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(gstat)
library(sp)
library(lubridate)
library(viridisLite)
```

## Import Excel data
```{r}
setwd("C:/Users/Daniel Jackson/Desktop/R/Warehouse Environments Analysis/")


# Import Excel data
data <- read_excel("./Data/Wifi_Test_B_Farms.xlsx")

# Look at the first rows
head(data)
```

## Convert the timestamp
```{r}
data <- data %>%
  mutate(
    Timestamp = as.POSIXct(
      Timestamp,
      format = "%Y-%m-%d %H:%M",
      tz = "UTC"
    )
  )

head(data$Timestamp)
```

# Define the room dimensions

For now, let's use the dimensions represented by our sample dataset:
```{r}
# ============================================================
# DEFINE ROOM DIMENSIONS
# ============================================================

# Coordinate system:
#
# Origin (0, 0, 0) = back-left-floor
#
# X = Width
# X increases from left → right
# Total width = 25 ft
#
# Y = Length
# Y increases from back → front
# Total length = 35 ft
#
# Z = Height
# Z increases from floor → ceiling
# Total height = 30 ft

room_width  <- 25
room_length <- 35
room_height <- 30


room <- list(
  width = room_width,
  length = room_length,
  height = room_height
)

room
```



## Create the room boundaries
```{r}
# ============================================================
# CREATE ROOM BOUNDARIES
# ============================================================

corners <- data.frame(

  X = c(
    0,
    room_width,
    room_width,
    0,
    0,
    room_width,
    room_width,
    0
  ),

  Y = c(
    0,
    0,
    room_length,
    room_length,
    0,
    0,
    room_length,
    room_length
  ),

  Z = c(
    0,
    0,
    0,
    0,
    room_height,
    room_height,
    room_height,
    room_height
  )

)

corners$ID <- c(
  "A",
  "B",
  "C",
  "D",
  "E",
  "F",
  "G",
  "H"
)

corners
```
The corners are:

          E -------- F
         /|         /|
        / |        / |
       G -------- H  |
       |  |       |  |
       |  A ------|--B
       | /        | /
       |/         |/
       C -------- D

Where:

A = bottom-back-left
B = bottom-back-right
C = bottom-front-right
D = bottom-front-left
E = top-back-left
F = top-back-right
G = top-front-right
H = top-front-left


Create the 12 edges
```{r}
# --------------------------------------------------
# Define the 12 edges
# --------------------------------------------------

edge_pairs <- data.frame(

  from = c(
    "A", "B", "C", "D",
    "E", "F", "G", "H",
    "A", "B", "C", "D"
  ),

  to = c(
    "B", "C", "D", "A",
    "F", "G", "H", "E",
    "E", "F", "G", "H"
  )

)

edge_pairs
```

### Convert the edges into Plotly coordinates
```{r}
# --------------------------------------------------
# Build plotting data for the room edges
# --------------------------------------------------

room_edges <- do.call(
  rbind,
  lapply(
    1:nrow(edge_pairs),
    function(i) {

      start <- corners[
        corners$ID == edge_pairs$from[i],
      ]

      end <- corners[
        corners$ID == edge_pairs$to[i],
      ]

      data.frame(
        X = c(start$X, end$X, NA),
        Y = c(start$Y, end$Y, NA),
        Z = c(start$Z, end$Z, NA)
      )
    }
  )
)

rownames(room_edges) <- NULL

head(room_edges, 15)
```
## Plot the room
```{r}
plot_ly(
  data = room_edges,

  x = ~X,
  y = ~Y,
  z = ~Z,

  type = "scatter3d",
  mode = "lines",

  line = list(
    width = 5,
    color = "black"
  ),

  name = "Room"
) %>%

  layout(

    scene = list(

      xaxis = list(
  title = "Width (ft)",
  range = c(0, room_width)
),

yaxis = list(
  title = "Length (ft)",
  range = c(0, room_length)
),

zaxis = list(
  title = "Height (ft)",
  range = c(0, room_height)
),

      aspectmode = "data"
    ),

    title = "Cold Storage Room Geometry"
  )
```

# Plot the actual sensor locations in 3D
```{r}
  plot_ly(
  data,
  x = ~X_ft,
  y = ~Y_ft,
  z = ~Z_ft,
  type = "scatter3d",
  mode = "markers",
  color = ~Temperature_F,
  colors = viridis(100),
  text = ~paste(
    "Sensor:", Sensor_ID,
    "<br>X:", X_ft,
    "<br>Y:", Y_ft,
    "<br>Z:", Z_ft,
    "<br>Temperature:", round(Temperature_F, 2), "°F",
    "<br>RH:", round(Relative_Humidity_pct, 1), "%"
  ),
  hoverinfo = "text",
  marker = list(
    size = 7
  )
) %>%
  layout(
    scene = list(
      xaxis = list(title = "Width X (ft)"),
      yaxis = list(title = "Length Y (ft)"),
      zaxis = list(title = "Height Z (ft)"),
      aspectmode = "data"
    ),
    title = "3D Sensor Locations"
  )

```

### Check that the sensors are within the room boundaries
```{r}
sensor_check <- data %>%

  mutate(

    Inside_Room =
      X_ft >= 0 &
      X_ft <= room_width &

      Y_ft >= 0 &
      Y_ft <= room_length &

      Z_ft >= 0 &
      Z_ft <= room_height

  )

table(sensor_check$Inside_Room)
```
Output should equal number of lines in the data frame. If not, some sensors are outside the room boundaries.

## Combine the Room and Sensors into the same 3D plot
```{r}
# ============================================================
# 3D ROOM + SENSOR LOCATIONS
# ============================================================

sensor_text <- paste(
  "Sensor:", data$Sensor_ID,
  "<br>X:", data$X_ft, " ft",
  "<br>Y:", data$Y_ft, " ft",
  "<br>Z:", data$Z_ft, " ft",
  "<br>Temperature:",
  round(data$Temperature_F, 2), " °F",
  "<br>RH:",
  round(data$Relative_Humidity_pct, 1), " %"
)


P <- plot_ly(

  # ----------------------------------------------------------
  # SENSOR TRACE
  # ----------------------------------------------------------

  x = data$X_ft,
  y = data$Y_ft,
  z = data$Z_ft,

  type = "scatter3d",
  mode = "markers",

  color = data$Temperature_F,

  colors = viridis(100),

  text = sensor_text,

  hoverinfo = "text",

  marker = list(
    size = 7
  ),

  name = "Sensors"

) %>%

  # ----------------------------------------------------------
  # ROOM TRACE
  # ----------------------------------------------------------

  add_trace(

    x = room_edges$X,
    y = room_edges$Y,
    z = room_edges$Z,

    type = "scatter3d",
    mode = "lines",

    line = list(
      width = 5,
      color = "black"
    ),

    name = "Room",

    hoverinfo = "none",

    inherit = FALSE

  ) %>%

  # ----------------------------------------------------------
  # LAYOUT
  # ----------------------------------------------------------

  layout(

    scene = list(

      # X = WIDTH
      xaxis = list(
        title = "Width (ft)",
        range = c(
          0,
          room_width
        )
      ),

      # Y = LENGTH
      yaxis = list(
        title = "Length (ft)",
        range = c(
          0,
          room_length
        )
      ),

      # Z = HEIGHT
      zaxis = list(
        title = "Height (ft)",
        range = c(
          0,
          room_height
        )
      ),

      aspectmode = "data"

    ),

    title = "3D Sensor Locations"

  )


P
```

## Add a Temperature_C column to the data frame
```{r}
data <- data %>%
  mutate(
    Temperature_C =
      (Temperature_F - 32) * 5 / 9
  )
```


## Calculate dew point
```{r}
data <- data %>%
  mutate(
    gamma =
      log(Relative_Humidity_pct / 100) +
      (
        17.625 * Temperature_C
      ) /
      (
        243.04 + Temperature_C
      ),

    Dew_Point_C =
      243.04 * gamma /
      (
        17.625 - gamma
      )

  )
```

## Calculate dew point depression
```{r}
data <- data %>%

  mutate(

    Dew_Point_F =
      Dew_Point_C * 9 / 5 + 32,

    Dew_Point_Depression_F =
      Temperature_F - Dew_Point_F

  ) %>%

  select(
    -gamma
  )
```

### Check to confirm 
```{r}
names(data)
```
# Interpolation

## Set Grid Resolution 
```{r}
grid_resolution_x <- 2  # ft
grid_resolution_y <- 2   # ft
grid_resolution_z <- 2  # ft

idw_power <- 2
```

## Create a grid for interpolation
```{r}
interpolation_grid <- expand.grid(

  X_ft = seq(
    0,
    room_width,
    by = grid_resolution_x
  ),

  Y_ft = seq(
    0,
    room_length,
    by = grid_resolution_y
  ),

  Z_ft = seq(
    0,
    room_height,
    by = grid_resolution_z
  )

)

nrow(interpolation_grid)
```
## IDW Interpolation Function
```{r}
# ============================================================
# IDW INTERPOLATION FUNCTION
# ============================================================

interpolate_variable <- function(
    sensor_data,
    grid,
    variable,
    idp = 2
) {

  # ----------------------------------------------------------
  # Keep complete observations for this variable
  # ----------------------------------------------------------

  sensor_data <- sensor_data %>%

    filter(
      !is.na(.data[[variable]]),
      !is.na(X_ft),
      !is.na(Y_ft),
      !is.na(Z_ft)
    )


  # ----------------------------------------------------------
  # Convert sensor locations to SpatialPointsDataFrame
  # ----------------------------------------------------------

  coordinates(sensor_data) <-
    ~X_ft + Y_ft + Z_ft


  # ----------------------------------------------------------
  # Convert interpolation grid to SpatialPoints
  # ----------------------------------------------------------

  coordinates(grid) <-
    ~X_ft + Y_ft + Z_ft


  # ----------------------------------------------------------
  # IDW interpolation
  # ----------------------------------------------------------

  result <- gstat::idw(

    formula =
      as.formula(
        paste(
          variable,
          "~ 1"
        )
      ),

    locations = sensor_data,

    newdata = grid,

    idp = idp

  )


  # ----------------------------------------------------------
  # Return ordinary data frame
  # ----------------------------------------------------------

  result <- as.data.frame(result)


  # Rename predicted value
  names(result)[
    names(result) == "var1.pred"
  ] <- paste0(
    "Predicted_",
    variable
  )


  # Keep only required fields
  result %>%

    select(
      X_ft,
      Y_ft,
      Z_ft,
      starts_with("Predicted_")
    )

}
```

## Test Function Using the Earliest Timestamp
```{r}

test_time <- min(
  data$Timestamp
)

test_sensors <- data %>%
  filter(
    Timestamp == test_time
  )
nrow(test_sensors)
```
```{r}
temperature_test <- interpolate_variable(
  sensor_data = test_sensors,
  grid = interpolation_grid,
  variable = "Temperature_F",
  idp = idw_power
)

head(temperature_test)
```


## Interpolate all four variables
```{r}
# ============================================================
# INTERPOLATE ALL ENVIRONMENTAL VARIABLES
# ============================================================

temperature_test <- interpolate_variable(
  test_sensors,
  interpolation_grid,
  "Temperature_F",
  idw_power
)


rh_test <- interpolate_variable(
  test_sensors,
  interpolation_grid,
  "Relative_Humidity_pct",
  idw_power
)


dew_point_test <- interpolate_variable(
  test_sensors,
  interpolation_grid,
  "Dew_Point_F",
  idw_power
)


dpd_test <- interpolate_variable(
  test_sensors,
  interpolation_grid,
  "Dew_Point_Depression_F",
  idw_power
)
```
### Combine the interpolated results into a single data frame
```{r}
# ============================================================
# COMBINE INTERPOLATED VARIABLES
# ============================================================

test_field <- temperature_test %>%

  left_join(
    rh_test,
    by = c(
      "X_ft",
      "Y_ft",
      "Z_ft"
    )
  ) %>%

  left_join(
    dew_point_test,
    by = c(
      "X_ft",
      "Y_ft",
      "Z_ft"
    )
  ) %>%

  left_join(
    dpd_test,
    by = c(
      "X_ft",
      "Y_ft",
      "Z_ft"
    )
  ) %>%

  mutate(
    Timestamp = test_time
  )
```

```{r}
names(test_field)
```
## Visualize the interpolated temperature field in 3D
```{r}
# ============================================================
# 3D INTERPOLATED TEMPERATURE FIELD
# ============================================================

plot_ly(

  data = test_field,

  x = ~X_ft,
  y = ~Y_ft,
  z = ~Z_ft,

  color = ~Predicted_Temperature_F,

  colors = viridis(
    100
  ),

  type = "scatter3d",

  mode = "markers",

  marker = list(
    size = 8,
    opacity = 0.20,
    symbol = "square"
  ),

  hoverinfo = "none"

) %>%

  layout(

    scene = list(

      xaxis = list(
        title = "Width X (ft)",
        range = c(
          0,
          room_width
        )
      ),

      yaxis = list(
        title = "Length Y (ft)",
        range = c(
          0,
          room_length
        )
      ),

      zaxis = list(
        title = "Height Z (ft)",
        range = c(
          0,
          room_height
        )
      ),

      aspectmode = "data"

    ),

    title = paste(
      "Interpolated Temperature Field —",
      format(
        test_time,
        "%Y-%m-%d %H:%M"
      )
    )

  )
```

## Add Sensor Data 
```{r}
# ============================================================
# ADD ACTUAL SENSOR LOCATIONS
# ============================================================

sensor_plot_data <- test_sensors
```

# Example Final Plot Showing Interpolated Field and Sensor Locations
```{r}
P <- plot_ly(

  # ==========================================================
  # INTERPOLATED FIELD
  # ==========================================================

  data = test_field,

  x = ~X_ft,
  y = ~Y_ft,
  z = ~Z_ft,

  type = "scatter3d",
  mode = "markers",

  marker = list(

    size = 10,

    opacity = 0.15,

    symbol = "square",

    color = test_field$Predicted_Temperature_F,

    colorscale = "Viridis",

    colorbar = list(
      title = list(
        text = "Temperature (°F)"
      ),
      x = 1.12
    )

  ),

  hoverinfo = "none",

  name = "Estimated Values"

) %>%

  # ==========================================================
  # SENSOR LOCATIONS
  # ==========================================================

  add_trace(

    data = sensor_plot_data,

    x = ~X_ft,
    y = ~Y_ft,
    z = ~Z_ft,

    type = "scatter3d",
    mode = "markers",

    marker = list(
      size = 9,
      color = "black"
    ),

    text = ~paste(

      "<b>Sensor:",
      Sensor_ID,
      "</b><br>",

      "Length X (ft): ",
      X_ft,
      "<br>",

      "Width Y (ft): ",
      Y_ft,
      "<br>",

      "Height Z (ft): ",
      Z_ft,
      "<br>",

      "Temperature: ",
      round(Temperature_F, 2),
      " °F<br>",

      "RH: ",
      round(Relative_Humidity_pct, 1),
      " %<br>",

      "Dew Point: ",
      round(Dew_Point_F, 2),
      " °F<br>",

      "Dew Point Depression: ",
      round(Dew_Point_Depression_F, 2),
      " °F"

    ),

    hoverinfo = "text",

    name = "Sensors",

    inherit = FALSE

  ) %>%

  # ==========================================================
  # LAYOUT
  # ==========================================================

  layout(

    scene = list(

      xaxis = list(
        title = "Width (ft)",
        range = c(
          0,
          room_width
        )
      ),

      yaxis = list(
        title = "Length (ft)",
        range = c(
          0,
          room_length
        )
      ),

      zaxis = list(
        title = "Height (ft)",
        range = c(
          0,
          room_height
        )
      ),

      aspectmode = "data"

    ),

    # --------------------------------------------------------
    # TRACE LEGEND
    # --------------------------------------------------------

    legend = list(

      x = 0.01,

      y = 0.99,

      xanchor = "left",

      yanchor = "top"

    ),

    title = "3D Environmental Field"

  )

P
```

## Create a function to interpolate all four variables for a given timestamp
```{r}
# ============================================================
# INTERPOLATE ONE TIMEPOINT
# ============================================================

interpolate_timepoint <- function(

    timestamp,
    data,
    grid,
    idp = 2

) {

  # ----------------------------------------------------------
  # Sensor observations for this timestamp
  # ----------------------------------------------------------

  sensor_data <- data %>%

    filter(
      Timestamp == timestamp
    )


  # ----------------------------------------------------------
  # Interpolate temperature
  # ----------------------------------------------------------

  temperature <- interpolate_variable(

    sensor_data,
    grid,
    "Temperature_F",
    idp

  )


  # ----------------------------------------------------------
  # Interpolate relative humidity
  # ----------------------------------------------------------

  humidity <- interpolate_variable(

    sensor_data,
    grid,
    "Relative_Humidity_pct",
    idp

  )


  # ----------------------------------------------------------
  # Interpolate dew point
  # ----------------------------------------------------------

  dew_point <- interpolate_variable(

    sensor_data,
    grid,
    "Dew_Point_F",
    idp

  )


  # ----------------------------------------------------------
  # Interpolate dew point depression
  # ----------------------------------------------------------

  dpd <- interpolate_variable(

    sensor_data,
    grid,
    "Dew_Point_Depression_F",
    idp

  )


  # ----------------------------------------------------------
  # Combine the four interpolated fields
  # ----------------------------------------------------------

  result <- temperature %>%

    left_join(
      humidity,
      by = c(
        "X_ft",
        "Y_ft",
        "Z_ft"
      )
    ) %>%

    left_join(
      dew_point,
      by = c(
        "X_ft",
        "Y_ft",
        "Z_ft"
      )
    ) %>%

    left_join(
      dpd,
      by = c(
        "X_ft",
        "Y_ft",
        "Z_ft"
      )
    ) %>%

    mutate(
      Timestamp = timestamp
    )


  result
}
```

## Test the function for a specific timestamp
```{r}
test_time <- min(
  data$Timestamp
)

test_field <- interpolate_timepoint(

  timestamp = test_time,

  data = data,

  grid = interpolation_grid,

  idp = idw_power

)
```
```{r}
nrow(test_field)
names(test_field)
```
## Interpolate all timestamps in the dataset
Sort Unique Timepoints
```{r}
timepoints <- sort(
  unique(data$Timestamp)
)

length(timepoints)
```
Interpolate all timepoints
```{r,results='hide'}
# ============================================================
# INTERPOLATE ALL TIMEPOINTS
# ============================================================

all_interpolated_fields <- lapply(

  timepoints,

  function(t) {

    message(
      "Processing: ",
      format(t, "%Y-%m-%d %H:%M")
    )

    interpolate_timepoint(

      timestamp = t,

      data = data,

      grid = interpolation_grid,

      idp = idw_power

    )

  }

)
```
## Combine the Interpolations 
```{r}
processed_environmental_fields <- bind_rows(
  all_interpolated_fields
)

nrow(processed_environmental_fields)
```
```{r}
table(
  processed_environmental_fields$Timestamp
)
```

# Add metadata to the processed environmental fields
```{r}
experiment_id <- "Wifi_Test_July_22"
```

```{r}
# ============================================================
# FINAL PROCESSED DATASET
# ============================================================

processed_environmental_fields_final <-
  processed_environmental_fields %>%

  mutate(

    Experiment_ID = experiment_id,

    Grid_Resolution_X_ft = grid_resolution_x,

    Grid_Resolution_Y_ft = grid_resolution_y,

    Grid_Resolution_Z_ft = grid_resolution_z,

    IDW_Power = idw_power

  ) %>%

  select(

    Experiment_ID,

    Timestamp,

    X_ft,
    Y_ft,
    Z_ft,

    Predicted_Temperature_F,

    Predicted_Relative_Humidity_pct,

    Predicted_Dew_Point_F,

    Predicted_Dew_Point_Depression_F,

    Grid_Resolution_X_ft,
    Grid_Resolution_Y_ft,
    Grid_Resolution_Z_ft,

    IDW_Power

  )
```

```{r}
names(processed_environmental_fields_final)
head(processed_environmental_fields_final)
```

# Export data as CSV
## Export Interpolated Data
```{r}
# ============================================================
# EXPORT PROCESSED ENVIRONMENTAL FIELD
# ============================================================

experiment_id <- "Wifi_Test_July_22nd"

processed_environmental_fields_final <-
  processed_environmental_fields %>%

  mutate(
    Experiment_ID = experiment_id,

    Grid_Resolution_X_ft = grid_resolution_x,
    Grid_Resolution_Y_ft = grid_resolution_y,
    Grid_Resolution_Z_ft = grid_resolution_z,

    IDW_Power = idw_power
  ) %>%

  select(
    Experiment_ID,
    Timestamp,

    X_ft,
    Y_ft,
    Z_ft,

    Predicted_Temperature_F,
    Predicted_Relative_Humidity_pct,
    Predicted_Dew_Point_F,
    Predicted_Dew_Point_Depression_F,

    Grid_Resolution_X_ft,
    Grid_Resolution_Y_ft,
    Grid_Resolution_Z_ft,

    IDW_Power
  )


write.csv(
  processed_environmental_fields_final,
  "./Processed_Data/Wifi_Test_July_22nd_processed_data.csv",
  row.names = FALSE
)
```

## Export Sensor Data for Visualization
```{r}
# ============================================================
# EXPORT SENSOR DATA
# ============================================================

processed_sensor_data <-
  data %>%

  mutate(
    Experiment_ID = experiment_id
  ) %>%

  select(
    Experiment_ID,

    Timestamp,

    Sensor_ID,

    X_ft,
    Y_ft,
    Z_ft,

    Temperature_F,
    Relative_Humidity_pct,

    Dew_Point_F,
    Dew_Point_Depression_F,

    Bin_ID
  )


write.csv(
  processed_sensor_data,
  "./Processed_Data/Wifi_Test_July_22nd_sensor.csv",
  row.names = FALSE
)
```

